---
title: 'OTUs: 97% vs 100%'
author: "Ross Cunning"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 3, digits = 9)
```

```{r setup_2, include=FALSE}
# Load all R scripts in "R/" directory
source("R/functions.R")
# Load package libraries
library(phyloseq); library(vegan); library(multcompView); library(reshape2); library(igraph); library(stringr)
# Set colors for plotting clades
taxcolors <- matrix(c("#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462"), 
                    dimnames=list(c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG")))
```

```{r}
# Load phyloseq objects
load("data/STJ2015_100_sym.RData"); phy100.f <- phy.f
load("data/STJ2015_sym.RData"); phy97.f <- phy.f

# Filter OTUs by minimum count
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97.f)[which(taxa_sums(phy97.f) >= n)]
taxa100 <- taxa_sums(phy100.f)[which(taxa_sums(phy100.f) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97.f)
phy100.f <- prune_taxa(names(taxa100), phy100.f)

# Filter samples by minimum count
# Set threshold number of reads
sn <- 100
# Remove samples with fewer reads than threshold
phy97.f <- prune_samples(sample_sums(phy97.f)>=sn, phy97.f)
phy100.f <- prune_samples(sample_sums(phy100.f)>=sn, phy100.f)

# Filter OTUs by minimum count again in case any dropped below threshold after filtering samples
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97.f)[which(taxa_sums(phy97.f) >= n)]
taxa100 <- taxa_sums(phy100.f)[which(taxa_sums(phy100.f) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97.f)
phy100.f <- prune_taxa(names(taxa100), phy100.f)

# Label clades and subtypes for filtered phyloseq object tax_tables
get.st <- function(df) {
  within(df, {
    Clade <- substr(hit, 1, 1)
    Subtype <- gsub(hit, pattern="_[A-Z]{2}[0-9]{6}", replacement="")
    Subtype <- gsub(Subtype, pattern="_multiple", replacement="")
    Subtype2 <- ifelse(as.numeric(sim)==100, paste0("'", Subtype, "'"),
                       paste0("'[", rep(rle(sort(Subtype))$values, times=rle(sort(Subtype))$lengths), "]'^", 
                              unlist(lapply(rle(sort(Subtype))$lengths, seq_len)))[order(order(Subtype))])
    #Subtype <- ifelse(as.numeric(sim)==100, Subtype, paste("*", Subtype, sep=""))
  })
}

tax_table(phy97.f) <- as.matrix(get.st(data.frame(tax_table(phy97.f), stringsAsFactors=FALSE)))
tax_table(phy100.f) <- as.matrix(get.st(data.frame(tax_table(phy100.f), stringsAsFactors=FALSE)))

```
* Minimum count to retain OTU: `r n`
* Minimum count to retain sample: `r sn`

#### Descriptive stats of Filtered datasets
```{r phy.f_histograms, fig.height=4, fig.width=8}
# Compute summary statistics
stats97.f <- data.frame(`97% OTUs`=t(phystats(phy97.f)), check.names=F)
stats100.f <- data.frame(`100% OTUs`=t(phystats(phy100.f)), check.names=F)

# Create and plot histograms
taxhist97 <- hist(log10(taxa_sums(phy97.f)), plot=F)
taxhist100 <- hist(log10(taxa_sums(phy100.f)), plot=F)
samhist97 <- hist(log10(sample_sums(phy97.f)), plot=F)
samhist100 <- hist(log10(sample_sums(phy100.f)), plot=F)

par(mfrow=c(2, 4), mar=c(3,3,1,1))
plot(taxhist97, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist100, col="black", main="100% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(samhist97, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist100, col="black", main="100% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)

# Create stats table
knitr::kable(cbind(stats97.f, stats100.f))
```


### Transform count data
Count data are transformed to both relative abundance (proportions) and square-root proportions for downstream statistical analyses.
```{r transform}
# Convert to proportion (relative abundance)
phy97.f.p <- transform_sample_counts(phy97.f, function(x) x/sum(x))
phy100.f.p <- transform_sample_counts(phy100.f, function(x) x/sum(x))
# Apply transformation function
transform <- function(x) sqrt(x/sum(x))  # Set transformation function
phy97.f.t <- transform_sample_counts(phy97.f, transform)  # Transform data
phy100.f.t <- transform_sample_counts(phy100.f, transform)
```

### Clade overview
```{r}
# Clade overview
cladeAbund <- aggregate(data.frame(RelAbund=rowSums(otu_table(phy97.f.p))),
                        by=list(Clade=data.frame(tax_table(phy97.f.p))$Clade), FUN=sum)
cladeAbund$Prop <- prop.table(cladeAbund$RelAbund)

bars <- barplot(cladeAbund$Prop*100, col=taxcolors, space=0,
                names.arg=cladeAbund$Clade, xlab=expression(paste(italic('Symbiodinium'), " Clade")),
                ylab="Relative abundance (%)")
text(bars, cladeAbund$Prop*100+2, labels=round(cladeAbund$Prop*100, 1), xpd=T)

cladeAbund$Notus <- table(data.frame(tax_table(phy97.f.p))$Clade)
cladeAbund

# Are Clade I's real?
CladeI <- subset_taxa(phy97.f.p, Clade=="I")
tax_table(CladeI)


# Clades by sample
source("~/Documents/Academia/HIMB/USVI/STJ2012/R/functions.R")
composition <- function(phy, col, legend=T) {
  samdat <- data.frame(sample_data(phy))
  #samdat$Genus <- factor(samdat$Genus, levels=rev(levels(samdat$Genus)))
  samdat$Species <- factor(samdat$Species, levels=rev(levels(samdat$Species)))
  samdat$Site <- factor(samdat$Location, levels=rev(levels(samdat$Location)))
  samdat <- samdat[with(samdat, order(Species, Site)), ]
  typerelabund <- as.matrix(otu_table(phy)[order(data.frame(tax_table(phy))$hit), 
                                           rownames(samdat)])
  sitebreaks <- c(as.character(samdat$Site), "X")==c("X", as.character(samdat$Site))
  sitebreaks <- which(sitebreaks==F) - 1
  spbreaks <- c(which(duplicated(samdat$Species)==F) - 1, nrow(samdat))
  
  
  # Make Barplot
  barplot(typerelabund, horiz=T, space=0, axes=F,axisnames=F, yaxs="i", col=col)
  rect(0, 0, par("usr")[2], par("usr")[4], lwd=1, xpd=T)
  axis(side=1, at=seq(0, 1, 0.1), line=0, tck=-0.025, mgp=c(0,0.25,0), cex.axis=0.7)
  mtext(side=1, "Relative abundance", cex=0.7, line=1)
  # Add legend
  if (legend==T) {
    legend(x=par("usr")[2]/2, y=par("usr")[4], xjust=0.5, yjust=0.25, horiz=T, bty="n", xpd=T, 
           cex=0.7, legend=c("A", "B", "C", "D", "F", "G"), fill=taxcolors, x.intersp=0.5)
    legend(x=par("usr")[2]*1.1, y=par("usr")[4]*0.75, xjust=0, yjust=0.1, bty="n", xpd=T, cex=1, 
           pt.cex=1, legend=c("White Point", "West Tektite", "Cocoloba", "Cabritte Horn", "Booby Rock"), fill=c("red", "orange", "yellow", "green", "blue"), y.intersp=0.7, 
           x.intersp=0.3)
  }
  # Add grouping bars for Site
  sitecolors <- matrix(c("red", "orange", "yellow", "green", "blue"), 
              dimnames=list(c("White Point", "West Tektite", "Cocoloba", "Cabritte Horn", "Booby Rock"))) 
  for (i in 1:length(sitebreaks)) {
    lines(c(0, 1), c(sitebreaks[i], sitebreaks[i]), lty=2, lwd=0.25)
    rect(1.01, sitebreaks[i], 1.04, sitebreaks[i+1], col=sitecolors[samdat$Site[sitebreaks[i]+1],], 
         lwd=0.25, xpd=T)
  }
  
  # Add lines to separate species and species names
  for (i in 1:length(spbreaks)) {
    lines(c(0, 1.07), c(spbreaks[i], spbreaks[i]), xpd=T, type="l", lwd=0.4)
    text(1.03, (spbreaks[i] + spbreaks[i+1]) / 2, xpd=T, pos=4, cex=0.8,
         labels=paste(samdat$Genus[which(duplicated(samdat$Species)==F)][i], "\n",
                      samdat$Species[which(duplicated(samdat$Species)==F)][i], sep=""))
  }
}

par(mfrow=c(1,1), mar=c(2, 1.5, 2, 10), lwd=0.1, cex=0.7, xpd=NA)
# Plot composition of 97% within-sample OTUs colored by clade
composition(phy97.f.p, col=taxcolors[factor(data.frame(tax_table(phy97.f.p))[order(data.frame(tax_table(phy97.f.p))$Subtype), ]$Clade, levels=c("A","B","C","D","F","G"))], legend=T)


composition(phy97.f.p, col=rainbow(ntaxa(phy97.f.p)), legend=F)
```

# *Symbiodinium* in each coral
For each coral species, barplots are presented showing the relative abundance of OTUs obtained by 100% and 97%-within-sample clustering. OTUs comprising more than 4% of a sample are labeled with the unique OTU number and the *Symbiodinium* subtype and NCBI GenBank accession number of the closest BLAST hit for that OTU in the reference database. OTU numbers and barplot colors are NOT comparable across clustering methods.

## *Pseudodiploria strigosa*  
```{r pstr, fig.height=8, fig.width=12}
source("R/functions.R")
# Create subsetted phyloseq objects for Pseudodiploria strigosa
pstr97.f.p <- subset_samples(phy97.f.p, Species=="Pstr")
pstr100.f.p <- subset_samples(phy100.f.p, Species=="Pstr")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(pstr97.f.p, main="97% OTUs")
otubarplot(pstr100.f.p, main="100% OTUs")
```

```{r pstr.net, fig.height=6, fig.width=6}
pstr.net <- makenet(pstr97.f.p, 0)
set.seed(54538)
plotnet(pstr.net)
```

## *Siderastrea siderea*  
```{r ssid, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
ssid97.f.p <- subset_samples(phy97.f.p, Species=="Ssid")
ssid100.f.p <- subset_samples(phy100.f.p, Species=="Ssid")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ssid97.f.p, main="97% OTUs")
otubarplot(ssid100.f.p, main="100% OTUs")
```

```{r ssid.net, fig.height=6, fig.width=6}
ssid.net <- makenet(ssid97.f.p, 0)
set.seed(54538)
plotnet(ssid.net)
```

## *Dendrogyra cylindrus*  
```{r dcyl, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
dcyl97.f.p <- subset_samples(phy97.f.p, Species=="Dcyl")
dcyl100.f.p <- subset_samples(phy100.f.p, Species=="Dcyl")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(dcyl97.f.p, main="97% OTUs")
otubarplot(dcyl100.f.p, main="100% OTUs")
```

```{r dcyl.net, fig.height=6, fig.width=6}
dcyl.net <- makenet(dcyl97.f.p, 0)
set.seed(54538)
plotnet(dcyl.net)
```

## *Favia fragum*  
```{r ffra, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
ffra97.f.p <- subset_samples(phy97.f.p, Species=="Ffra")
ffra100.f.p <- subset_samples(phy100.f.p, Species=="Ffra")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ffra97.f.p, main="97% OTUs")
otubarplot(ffra100.f.p, main="100% OTUs")
```

```{r ffra.net, fig.height=6, fig.width=6}
ffra.net <- makenet(ffra97.f.p, 0)
set.seed(54538)
plotnet(ffra.net)
```

## *Montastraea cavernosa*  
```{r mcav, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
mcav97.f.p <- subset_samples(phy97.f.p, Species=="Mcav")
mcav100.f.p <- subset_samples(phy100.f.p, Species=="Mcav")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(mcav97.f.p, main="97% OTUs")
otubarplot(mcav100.f.p, main="100% OTUs")
```

```{r mcav.net, fig.height=6, fig.width=6}
mcav.net <- makenet(mcav97.f.p, 0)
set.seed(54538)
plotnet(mcav.net)
```

## *Madracis mirabilis*  
```{r mmir, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
mmir97.f.p <- subset_samples(phy97.f.p, Species=="Mmir")
mmir100.f.p <- subset_samples(phy100.f.p, Species=="Mmir")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(mmir97.f.p, main="97% OTUs")
otubarplot(mmir100.f.p, main="100% OTUs")
```

```{r mmir.net, fig.height=6, fig.width=6}
mmir.net <- makenet(mmir97.f.p, 0)
set.seed(54538)
plotnet(mmir.net)
```

## *Orbicella annularis*  
```{r oann, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
oann97.f.p <- subset_samples(phy97.f.p, Species=="Oann")
oann100.f.p <- subset_samples(phy100.f.p, Species=="Oann")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(oann97.f.p, main="97% OTUs")
otubarplot(oann100.f.p, main="100% OTUs")
```

```{r oann.net, fig.height=6, fig.width=6}
oann.net <- makenet(oann97.f.p, 0)
set.seed(54538)
plotnet(oann.net)
```

## *Porites astreoides*  
```{r past, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
past97.f.p <- subset_samples(phy97.f.p, Species=="Past")
past100.f.p <- subset_samples(phy100.f.p, Species=="Past")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(past97.f.p, main="97% OTUs")
otubarplot(past100.f.p, main="100% OTUs")
```

```{r past.net, fig.height=6, fig.width=6}
past.net <- makenet(past97.f.p, 0)
set.seed(54538)
plotnet(past.net)
```

## *Siderastrea radians*  
```{r srad, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
srad97.f.p <- subset_samples(phy97.f.p, Species=="Srad")
srad100.f.p <- subset_samples(phy100.f.p, Species=="Srad")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(srad97.f.p, main="97% OTUs")
otubarplot(srad100.f.p, main="100% OTUs")
```

```{r srad.net, fig.height=6, fig.width=6}
srad.net <- makenet(srad97.f.p, 0)
set.seed(54538)
plotnet(srad.net)
```

## *Water*  
```{r watr, fig.height=8, fig.width=12}
# Create subsetted phyloseq objects for Pseudodiploria strigosa
watr97.f.p <- subset_samples(phy97.f.p, Species=="Water")
watr100.f.p <- subset_samples(phy100.f.p, Species=="Water")
# Plot custom barplots for Pseudodiploria strigosa
par(mfrow=c(2,1), mar=c(4,4,2,2), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(watr97.f.p, main="97% OTUs")
otubarplot(watr100.f.p, main="100% OTUs")
```

```{r watr.net, fig.height=6, fig.width=6}
watr.net <- makenet(watr97.f.p, 0)
set.seed(54538)
plotnet(watr.net)
```


